### 2024-06-12 code/processing-code/processingfile.qmd
created processing file to start data cleaning and exploration. Added codebook for reference. renamed data file to be in line with naming conventions for Repo.

---
title: "An example cleaning script"
author: "Andreas Handel"
date: "2024-02-07"
output: html_document
editor: 
  markdown: 
    wrap: sentence
---

# Processing script

### Setup

Load needed packages.\
*Pacman::p_load* makes sure they are installed before auto-loading.\
*tidyverse* for general data processing/cleaning.\
*lubridate* for cleaning/re-coding date fields.\
*skimr* for nice visualization of data.\
*here* to set paths.

```{r}
#install.packages(pacman)
pacman::p_load(tidyverse,lubridate,skimr,here)
```

# Data loading

```{r}
# path to data
# note the use of the here() package and not absolute paths
data_location <- here::here("data","raw-data","raw_crime_060424.csv")
rawdata <- read.csv(data_location, TRUE)
```

# Check data

First we can look at the codebook, which can be found [here](https://data.austintexas.gov/Public-Safety/Crime-Reports/fdj4-gpfu/about_data).
I will recreate it here as it currently exists, and I will have a copy of the pdf attachment codebook included in the repo as well.
However, please note if this analysis is ever repeated that careful consideration should be taken to ensure field have not been changed, removed, or redefined.

| Field                       | Description                                                                                                  |
|------------------------------|------------------------------------------|
| Incident Number             | Incident report number                                                                                       |
| Highest Offense Description | Description                                                                                                  |
| Highest Offense Code        | Code                                                                                                         |
| Family Violence             | Incident involves family violence? Y = yes, N = no                                                           |
| Occurred Date Time          | Date and time (combined) incident occurred                                                                   |
| Occurred Date               | Date the incident occurred                                                                                   |
| Occurred Time               | Time the incident occurred                                                                                   |
| Report Date Time            | Date and time (combined) incident was reported                                                               |
| Report Date                 | Date the incident was reported                                                                               |
| Report Time                 | Time the incident was reported                                                                               |
| Location Type               | General description of the premise where the incident occurred                                               |
| Address                     | Incident location                                                                                            |
| Zip code                    | Zip code where incident occurred                                                                             |
| Council District            | Austin city council district where the incident occurred                                                     |
| APD Sector                  | APD sector where incident occurred                                                                           |
| APD District                | APD district where incident occurred                                                                         |
| PRA                         | APD police reporting area where incident occurred                                                            |
| Census Tract                | Census tract where incident occurred                                                                         |
| Clearance Status            | How/whether crime was solved (see Clearance lookup)                                                          |
| Clearance Date              | Date crime was solved                                                                                        |
| UCR Category                | Code for the most serious crimes identified by the FBI as part of its Uniform Crime Reporting program        |
| Category Description        | Description for the most serious crimes identified by the FBI as part of its Uniform Crime Reporting program |
| X-coordinate                | X-coordinate where the incident occurred                                                                     |
| Y-coordinate                | Y-coordinate where incident occurred                                                                         |
| Latitude                    | Latitude where incident occurred                                                                             |
| Longitude                   | Longitude where the incident occurred                                                                        |
| Location                    | 3rd party generated spatial column                                                                           |

| Clearance lookup |                      |
|------------------|----------------------|
| C                | Cleared by Arrest    |
| O                | Cleared by Exception |
| N                | Not cleared          |

Several ways of looking at the data

```{r}
dplyr::glimpse(rawdata)
summary(rawdata)
head(rawdata)
skimr::skim(rawdata)
unique(rawdata$Location.Type)
```

# Cleaning

By inspecting the data as done above, we find some problems that need addressing:

```{r}
d1 <- rawdata %>% dplyr::filter( Height != "sixty" ) %>% 
                  dplyr::mutate(Height = as.numeric(Height))
skimr::skim(d1)
hist(d1$Height)
```

Now we see that there is one person with a height of 6.
That could be a typo, or someone mistakenly entered their height in feet.
Since we unfortunately don't know, we might need to remove this person, which we'll do here.

```{r}
d2 <- d1 %>% dplyr::mutate( Height = replace(Height, Height=="6",round(6*30.48,0)) )
skimr::skim(d2)
```

Height values seem ok now.

Now let's look at the `Weight` variable.
There is a person with weight of 7000, which is impossible, and one person with missing weight.
To be able to analyze the data, we'll remove those individuals as well.

```{r}
d3 <- d2 %>%  dplyr::filter(Weight != 7000) %>% tidyr::drop_na()
skimr::skim(d3)
```

Now checking the `Gender` variable.
Gender should be a categorical/factor variable but is loaded as character.
We can fix that with simple base R code to mix things up.

```{r}
d3$Gender <- as.factor(d3$Gender)  
skimr::skim(d3)
```

Now we see that there is another NA, but it's not `NA` from R, instead it was loaded as character and is now considered as a category.
Well proceed here by removing that individual with that NA entry.
Since this keeps an empty category for Gender, I'm also using droplevels() to get rid of it.

```{r}
d4 <- d3 %>% dplyr::filter( !(Gender %in% c("NA","N")) ) %>% droplevels()
skimr::skim(d4)
```

All done, data is clean now.

Let's assign at the end to some final variable, this makes it easier to add further cleaning steps above.

```{r}
processeddata <- d4
```

# Save data

Finally, we save the clean data as RDS file.
I suggest you save your processed and cleaned data as RDS or RDA/Rdata files.
This preserves coding like factors, characters, numeric, etc.
If you save as CSV, that information would get lost.
However, CSV is better for sharing with others since it's plain text.
If you do CSV, you might want to write down somewhere what each variable is.

See here for some suggestions on how to store your processed data: http://www.sthda.com/english/wiki/saving-data-into-r-data-format-rds-and-rdata

```{r}
save_data_location <- here::here("data","processed-data","processeddata.rds")
saveRDS(processeddata, file = save_data_location)
```

# Notes

Removing anyone who had "faulty" or missing data is one approach.
It's often not the best.
based on your question and your analysis approach, you might want to do cleaning differently (e.g. keep individuals with some missing information).