### 2024-06-14 code/eda-code/eda.qmd
Significant amount more cleaning done. Created a processed data file and saved to data folder. Removed several template files.

version https://git-lfs.github.com/spec/v1
oid sha256:a063832bfc7ed6b54f7db32a75d484bdb6f59df9d1fba1d7ca6f90fe7911162f
size 4505