### 2024-06-14 products/manuscript/Manuscript.qmd

version https://git-lfs.github.com/spec/v1
oid sha256:53791af84dddf4d0e7fd1fcad04d4f30f60de0ca33f13bf28575f0f87d12c65f
size 4288