### 2024-08-03 code/analysis-code/analysisfile.qmd
Project Assignment 4 and 5 submission. Mostly ready for final submission, need to add narrative to eda and stat qmd files

---
title: "Analysis"
author: "William Hyltin"
date: today
output: html_document
editor: 
  markdown: 
    wrap: sentence
---

# Setup

```{r setup}
#load needed packages. make sure they are installed.
pacman::p_load(here, knitr, tidyverse, skimr, fpp3, urca, forecast, gt)
theme_set(theme_minimal())
```

```{r}
d1 <- readRDS(here('data','processed-data','processed-crime.rds'))
```

```{r sandbox}

d1 %>% mutate(Category.Description = if_else(Category.Description == '', 'NA/Unknown', Category.Description)) %>% group_by(Category.Description, Highest.Offense.Description) %>% summarize(cnt = n()) %>% pivot_wider(names_from = Category.Description, values_from = cnt)

temp <- d1 %>%
         mutate(Occurred.Date = trunc.Date(Occurred.Date, 'months')) %>% 
         group_by(Occurred.Date) %>% 
         summarize(
             cnt = n()
         ) %>% as.data.frame() %>% 
  filter(year(Occurred.Date) != 2024)
rownames(temp) <- temp$Occurred.Date
temp <- temp %>% select(cnt)

temp2 <- temp %>% as.ts()
Acf(temp2)
Box.test(temp2, lag = 12, type = 'Ljung-Box')
auto.arima(temp2)
auto.arima(temp2) %>% forecast(h=24) %>% autoplot()
#temp2 %>% log() %>% diff(12) %>% diff() %>% diff() %>% diff() %>% diff() %>% diff() %>% Acf()
```

```{r}
temp3 <- d1 %>%
         filter(Crime.Category == 'MURDER') %>% 
         mutate(Occurred.Date = trunc.Date(Occurred.Date, 'months')) %>% 
         group_by(Occurred.Date) %>% 
         summarize(
             cnt = n()
         ) %>% as.data.frame() %>% 
  filter(year(Occurred.Date) < 2024)
rownames(temp3) <- temp3$Occurred.Date
temp3 <- temp3 %>% select(cnt)

temp5 <- temp3 %>% as.ts()
Acf(temp5)
auto.arima(log(temp5))
Box.test(diff(log(temp5)), lag = 4, type = 'Ljung-Box')
temp5 %>% log() %>% diff() %>% ur.kpss() %>% summary()

temp5 %>% log() %>% diff() %>% Acf()
temp5 %>% log() %>% diff() %>% plot()
```

```{r}
fit <- Arima(temp5, order = c(0,1,1))
autoplot(forecast(fit, h=24))
fit2 <- Arima(log(temp5), order = c(0,1,1))
autoplot(forecast(fit2, h=24))
```

```{r}
fets <- function(x, h) {
  forecast(ets(x), h=h)
}
farima <- function(x, h) {
  forecast(auto.arima(x), h=h)
}

e1 <- tsCV(log(temp5), fets, h=1)
e2 <- tsCV(log(temp5), farima, h=1)

# MSE comparisons
mean(e1^2, na.rm=TRUE)
mean(e2^2, na.rm=TRUE)
```

```{r}
crime_agg <- d1 %>%
  filter(Occurred.Date <= '2024-03-31') %>% 
  mutate(Occurred.Date = yearmonth(Occurred.Date)) %>% 
  group_by(Occurred.Date, Crime.Category) %>% 
  summarize(
     cnt = n()
  ) %>% as.data.frame() %>% 
  as_tsibble(key = Crime.Category, index = Occurred.Date) %>% 
  aggregate_key(Crime.Category, Count = sum(cnt)) %>% 
  fill_gaps(Count = 0)

group_fit <- crime_agg %>% 
  filter(year(Occurred.Date) < 2024) %>%
  model(base = ARIMA(Count)) %>% 
  reconcile(
    td = top_down(base)
  )

group_fit

group_fcast <- group_fit %>% 
  forecast(h = '2 years')

grpPlot <- group_fcast %>% 
  filter(Crime.Category %in% c('MURDER', 'AGGRAVATED ASSAULT', 'UNLAWFUL RESTRAINT' )| is_aggregated(Crime.Category), .model == 'td') %>% 
  autoplot(
    crime_agg %>% filter(year(Occurred.Date) >= 2019)#, Crime.Category == 'MURDER')
  ) + facet_wrap(vars(Crime.Category), scales = 'free_y',
                 labeller = labeller(
                   Crime.Category = c('<aggregated>' = 'All Crime', 
                                      'AGGRAVATED ASSAULT' = 'Aggravated Assault', 
                                      'MURDER' = 'Murder', 
                                      'UNLAWFUL RESTRAINT' = 'Unlawful Restraint')
                   )
                 ) +
  labs(x = 'Month', y = 'Count', title = 'Grouped Forecasts of Crime Occurence by Month') +
  theme(panel.background = element_rect(fill = 'white',
                                        color = 'white'),
        plot.background = element_rect(fill = 'white',
                                        color = 'white')
        )

murdplot <- group_fcast %>% 
  filter(Crime.Category %in% c('MURDER'), .model == 'td') %>% 
  autoplot(
    crime_agg %>% filter(year(Occurred.Date) >= 2019, Crime.Category == 'MURDER')
  ) +
  labs(x = 'Month', y = 'Count', title = 'Murder Forecasts by Month') +
  theme(panel.background = element_rect(fill = 'white',
                                        color = 'white'),
        plot.background = element_rect(fill = 'white',
                                        color = 'white')
        )

grpPlot  
murdplot

ggsave(here('results', 'figures', 'static-plots', 'crime-forecast.png'), grpPlot)
ggsave(here('results', 'figures', 'static-plots', 'murder-forecast.png'), murdplot)
```
```{r}
crime_agg %>% 
    filter(year(Occurred.Date) < 2024,Crime.Category == 'MURDER') %>%
    model(base = ARIMA(Count))

crime_agg %>% 
    filter(year(Occurred.Date) < 2024,Crime.Category == 'MURDER') %>%
    model(base = ARIMA(Count)) %>% forecast(h = '2 years') %>% 
  autoplot(crime_agg %>% filter(year(Occurred.Date) >= 2019, Crime.Category == 'MURDER'))
```

```{r}
AvF <- d1 %>% 
  filter(year(Occurred.Date) == 2024, Crime.Category == 'MURDER') %>% 
  mutate(
    Occurred.Date = yearmonth(Occurred.Date)
    ) %>% 
  group_by(Occurred.Date) %>% 
  summarize(
    n = n()
    ) %>% left_join(filter(group_fcast, Crime.Category == 'MURDER' & .model == 'td'), join_by(Occurred.Date == Occurred.Date)) %>% 
  select(Occurred.Date, n, .mean) %>% 
  gt() %>% 
  tab_header(
    title = md('**Forecast vs Actual<br>Murder Count**')
    ) %>% 
  cols_label(
    Occurred.Date = md('**Month**'),
    n = md('**Actual**'),
    .mean = md('**Forecasted**')
  ) %>% 
  fmt_number(columns = .mean, decimals = 1)

AvF

gtsave(AvF, 'actual-vs-forecast.html', here('results', 'figures', 'static-plots'))
```

```{r}
aggresid <- group_fit %>% filter(is_aggregated(Crime.Category)) %>% select(td) %>% gg_tsresiduals()

aggresid
ggsave(here('results', 'figures', 'static-plots', 'crime-resid.png'), aggresid)
```

```{r}
murdresid <- group_fit %>% filter(Crime.Category == 'MURDER') %>% select(td) %>% gg_tsresiduals()

murdresid
ggsave(here('results', 'figures', 'static-plots', 'murd-resid.png'), murdresid)
```

```{r}
agglb <- augment(filter(group_fit, is_aggregated(Crime.Category))) %>% 
  features(.resid, ljung_box, l = 24) %>% gt()

agglb
gtsave(agglb, 'aggregated-ljung.html', here('results', 'figures', 'static-plots'))
```

```{r}
murdlb <- augment(filter(group_fit, Crime.Category == 'MURDER')) %>% 
  features(.resid, ljung_box, l = 24) %>% gt()

murdlb
gtsave(murdlb, 'murder-ljung.html', here('results', 'figures', 'static-plots'))
```